---
title: "STAT/MATH 495: Problem Set 07"
author: "Jenn Halbleib"
date: "2017-10-24"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
library(ggplot2)
train <- read_csv("data/cs-training.csv") %>% 
  rename(Id = X1)
test <- read_csv("data/cs-test.csv") %>% 
  rename(Id = X1)
submission <- read_csv("data/sampleEntry.csv")
```

Information on the competition can be found [here](https://www.kaggle.com/c/GiveMeSomeCredit/data).



# Collaboration

Please indicate who you collaborated with on this assignment:  
*Quick Start Guide from course webpage

#Exploratory Data Analysis

```{r}
#Distribution of outcome variable SeriousDlqin2yrs
train <- train %>% mutate(cat_deliq = ifelse( SeriousDlqin2yrs == 1, "yes", "no"))
ggplot(data=train, aes(x = cat_deliq)) + geom_bar() + ggtitle("Distribution of Serious Deliquency") + xlab("Serious Deliquency")

#Building boxplots of DebtRatio, age, and MontlyIncome against SeriousDlqin2yrs

#Adding 0.00001 to DebtRatio to enable taking the log (many zeros in DebtRatio)
train <- train %>% mutate(log_debt_ratio = log(DebtRatio + 0.00001))

#DebtRatio boxplot
ggplot(data=train,aes(cat_deliq,log_debt_ratio)) + geom_boxplot() + xlab("Serious Deliquency") + ylab("Natural Log of Debt Ratio") + ggtitle("Boxplot of Serious Deliquency to Debt Ratio") + coord_flip()

#age boxplot
ggplot(data=train,aes(cat_deliq,age)) + geom_boxplot() + xlab("Serious Deliquency") + ylab("Age") + ggtitle("Boxplot of Serious Deliquency to Age") + coord_flip()

#Adding 0.00001 to Monthly Income to enable taking the log
train <- train %>% mutate(log_monthly_income = log(MonthlyIncome + 0.00001))

#MonthlyIncome boxplot, generates warning message for rows with NA values
ggplot(data=train,aes(cat_deliq,log_monthly_income)) + geom_boxplot() + xlab("Serious Deliquency") + ylab("Natural Log of Monthly Income") + ggtitle("Boxplot of Serious Deliquency to Monthly Income") + coord_flip()
```

# Build binary classifier

Examining the 3 boxplots, age appears to have the largest mean difference with serious deliquency. For this reason, I've chosen age as the predictor.

```{r}
model_formula <- as.formula(SeriousDlqin2yrs~age)
model_logistic <- glm(model_formula, data=train, family="binomial")

#Extract regression table in tidy format
model_logistic %>% 
  broom::tidy(conf.int=TRUE)
```

# ROC curve

Based on the ultimate classifier you choose, plot a corresponding ROC curve.

```{r}
profiles_train_augmented <- model_logistic %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1+exp(-.fitted)))

library(ROCR)
# This bit of code computes the ROC curve
pred <- prediction(predictions = profiles_train_augmented$p_hat, labels = profiles_train_augmented$SeriousDlqin2yrs)
perf <- performance(pred, "tpr","fpr")

# This bit of code computes the Area Under the Curve
auc <- as.numeric(performance(pred,"auc")@y.values)
auc

# This bit of code prints it
plot(perf, main=paste("Area Under the Curve =", round(auc, 3)))
abline(c(0, 1), lty=2)
```

#Making submission file
```{r}
final <- predict(model_logistic,test) 
final <- final %>% as_tibble() %>% mutate(Probability = 1/(1+exp(-value))) 
final <- final %>% rownames_to_column(var = "ID")
final_submission <- final %>% select(ID,Probability)
write_csv(final_submission, "submission.csv")
```


# ROC curve for random guessing

Instead of using any predictor information as you did above, switch your
predictions to random guesses and plot the resulting ROC curve.

```{r}
#Generating random guesses
random_guesses <- rbinom(150000, 1, 0.5)
pred2 <- prediction(predictions = random_guesses, labels = profiles_train_augmented$SeriousDlqin2yrs)
perf <- performance(pred2, "tpr","fpr")

# This bit of code computes the Area Under the Curve
auc2 <- as.numeric(performance(pred2,"auc")@y.values)
auc2

# This bit of code prints it
plot(perf, main=paste("Area Under the Curve =", round(auc2, 3)))
abline(c(0, 1), lty=2)
```